---
title: "Mini-Project 2"
author: "Danica Miguel, Sophia Tannir, and Emily Daubenspeck"
date: "23 March 2018"
output:
  html_document:
    code_folding: hide
---

## Background 

#### In the age of Big Data, individuals are demonstrating a proclivity towards data-driven tools in order to glean insight into their chances at success in certain areas. Nowhere is this seen as prominently as in the world of higher education, where students are free to use online services like [Niche](https://www.niche.com/colleges/admissions-calculator/) to plot themselves as data points against prior applicants to colleges; typically, individuals provide two numeric values, usually a score and/or GPA, which act as the axes for the generated data visualization. In their most basic form, these plots help determine if admission to a given school is likely. **With our data visualization^[[GitHub Repository](https://github.com/ed4ubenspeck/sds192-mp2.git)], we have created a similar experience for the aspiring politician; where do a candidate’s campaign finance metrics place them among past candidates, and are they on the path of those who have won, or those who have lost?**



```{r, include=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
library(scales)
library(gridExtra)
```


```{r, message = FALSE, warning = FALSE}
# joining tables to align candidate IDs with relevant information on campaign contributions
con_can <- candidates %>%
  inner_join(contributions, by = "cand_id") %>%
  select(fec_id = cand_id, 
         cand_name, 
         cand_party_affiliation, 
         cand_election_yr, 
         cand_state,
         cmte_id, 
         committee_name = name,
         transaction_amt, 
         tran_id) %>%
  # ommission of rows with missing data
  na.omit()
```


```{r fig.width = 20, fig.height = 10, message = FALSE, warning = FALSE}

# function to further refine table by state
cut_down_table <- function(state_arg) {
  con_can %>%
    filter(cand_state == state_arg) %>%
    group_by(fec_id, 
             cand_election_yr) %>%
    # summary of total money received by candidate, as well as total number of committees contributing per candidate
    summarize(money_sum = sum(transaction_amt), 
              total_com = n_distinct(committee_name))
}

# function to join with house_elections to provide information on election year and result
join_for_wins <- function(state_arg) {
  cut_down_table(state_arg) %>%
    inner_join(house_elections, by = "fec_id") %>%
    select(fec_id, 
           money_sum, 
           total_com, 
           state,
           candidate_name, 
           party, 
           ge_winner)
}

# function to generate individual plots
state_plot <- function(state_arg) {
  join_for_wins(state_arg) %>%
    ggplot(aes(x = total_com, 
               y = money_sum, 
               fill = ge_winner)) +
    geom_point(position = position_jitter(width = 1, height = 0.5), 
               shape = 21, 
               color = "transparent", 
               size = 3) +
    scale_fill_manual(values = c(W = "orange", N = "blue"), 
                      label = c(W = "Win", N = "No Win")) +
    stat_smooth(method = lm, 
                se = TRUE, 
                fullrange = TRUE,
                alpha = .2,
                aes(fill = ge_winner, color = ge_winner), 
                show.legend = FALSE) +
    ggtitle(state_arg) +
    theme(plot.title = element_text(size = 40, face = "bold"), legend.text = element_text(size = 15.5)) +
    scale_y_continuous(labels = dollar) +
    ylab("Total Funds Received") +
    theme(axis.title.y = element_text(margin = margin(r = 30), size = 30)) +
    theme(axis.text.y = element_text(size = 20)) +
    xlab("Total Committees Donating") +
    theme(axis.title.x = element_text(margin = margin(t = 30), size = 30)) +
    theme(axis.text.x = element_text(margin = margin(t = 20), size = 20)) +
    scale_x_continuous(limits = c(0, 150)) +
    theme(legend.title = element_blank()) +
    scale_color_manual(values = c(W = "orange", N = "blue"))
}

# defining populous states to plot for
populous_states <- c(
  "TX",
  "CA",
  "NY",
  "PA",
  "IL",
  "OH"
)

# making a list of all plots
swing_grid <- lapply(populous_states, state_plot)

# arranging all plots into a grid for viewing
do.call("grid.arrange", c(swing_grid, ncol = 3))
```

## Examining the Graphic

#### Concerning variable selection, we wanted to provide insight into the total amount of committee money that was received in contributions, but we were torn between plotting this against either the total number of committee transactions or the total number of committees. Eventually, we chose the latter; it is possible that this value could indicate the breadth of support for a candidate when comparing candidates who received funding from only a few committees to those who received contributions from many. 

#### Say you're a candidate from California. You want to know where your current campaign financial metric puts you relative to previous winning candidates. You estimate that you'll receive backing from approximately 75 seperate committees, but you're unsure of the total funding you'll need to win. Looking at the California plot, you might see that to fall within the confidence interval for winning candidates, you will need to raise at least 2.5 million dollars to win. This graphic can therefore work to heighten the efficiency of candidate fundraising and further promote the role of predictive statistics in the formulation of individual campaign efforts. 

## Final thoughts

#### Due to the nature of the data provided, we were limited in what we could produce while maintaining statistical validity. We chose not to facet by political party, as this resulted in plots with very few points. Furthermore, we chose to limit our work to the six most populous states in the U.S, as they provided much more data, as more populous states hold more seats in the House of Representatives. This allowed us to create something more in line with our original vision for the graphic. Finally, we opted not to omit incumbents who were running unopposed, as this would have limited our pool of data even further. Rather than working to perform a deeper analysis of this data as a whole, we were most interested in working to produce something evocative of one of today’s most ubiquitous, interactive data visualizations. With our data graphic, a prospective candidate in any of the six states considered would be able to examine her campaign finance metrics and plot herself against those who have come before her; she might pay special attention to her own point’s position relative to the Win or No Win confidence intervals, and perhaps adjust her financing goals accordingly. 


Current word count: 545


